---
title: "Thesis Draft : Art Auction Data"
name: "Joey Yoo"
output: html_notebook
---

```{r global_options, include=FALSE}
# Set the working directory for current step
setwd("/Users/maeror/Documents/projects/contemporary_art_thesis/")
# Load the libraries
library(xtable)
library(Matrix)
library(lme4)
library(numDeriv)
library(dplyr)
library(ggplot2)
library(knitr)
library(lme4)
library(readxl)
library(stringdist)
library(data.table)
library(utils)
library(gridExtra)
library(RLRsim)
# Load the functions from src file
# source("src/algorithm3.R")
# source("src/algorithm3BootInt.R")
# source("src/auxFunctions.R")
# source("src/intervalStr.R")
# source("src/algorithm3SEWorking.R")
# #source("algorithm3SE.R")
opts_chunk$set(tidy.opts=list(width.cutoff=50), tidy=TRUE, cache = TRUE)
```

## RAW DATA CLEANSING WITH R
```{r}
# Read the excel sheets and merge them to a single dataframe
read_excel_allsheets <- function(filename, tibble = FALSE, bysheets = FALSE) {
  sheets <- readxl::excel_sheets(filename)
  x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
  if(!tibble) x <- lapply(x, as.data.frame)
  names(x) <- sheets
  if (bysheets) {
    return(x)
  } else {
    return(rbindlist(x))
  }
}
```

```{r}
### Artworks with more than 1 artists are ommited (15 accounts)
art1019 = read_excel_allsheets("data/artdata1019.xlsx")
# Omit title for now
data = art1019[,-2]
# Change column types
data$Birth = as.numeric(data$Birth)
data$Death = as.numeric(data$Death)
# Omit Birth NAs (artworks with multiple artists)
data = data[!is.na(data$Birth),]
# Categorial Variable for death
data = cbind(data, "dead" = factor(ifelse(is.na(data$Death), 0, 1)))
# Turn Location into Factor
data$Loc = factor(data$Loc)
# Add logrise
geomrise = log(data$Price / sqrt(data$Low*data$High))
data = cbind(data, "geomrise" = geomrise)
# Filter only the contemporary artists (1900~ 2000)
data = data[between(data$Birth,1900, 2000),]
# Separate Time
data$Year = as.numeric(substr(data$Date,1,4))

head(data)
write.csv(data, "art_mod1.csv")
# Modify art_mod1 with python
```

### DATA LOADING
```{r}
#### LOAD DATA HERE!!!! #####
# ARTDATA_MOD2 saves the record with all typo and birth year corrected.
data = read.csv("art_mod2.csv", header=TRUE,stringsAsFactors = FALSE)
# remove non-char data (wrong names)
data = data[,-1]
data = data[!grepl(":", data$Artist),]
data = data[!is.na(data$Artist),]

data$young = as.factor(data$young)
data = data[,-11]
# Unite the artist names, using the string distance
## Unite to the most occuring name
# originally 4057 of unique artists
data = cbind(data, "dead" = factor(ifelse(is.na(data$Death), 0, 1)))
data$age = data$Year - data$Birth 
data = data[!is.na(data$age),]
data$young = data$dead
levels(data$young) = c("0","3")
data$young = as.character(data$young)
data[data$young == 0,]$young = as.character(factor(cut(data[data$dead==0,]$age, 3, labels=c(0,1,2))))
data$young = as.factor(data$young)
```


```{r}
# Augment the Data with currency information
currUSD = read.csv("./1019 Currency_against_USD - Sheet1.csv")
Rate = rep(0, dim(data)[1])

for (i in 1:dim(data)[1]) {
  curr_date = data[i,]$Date
  currency = data[i,]$Curr
  year = as.integer(substr(curr_date,1,4))
  month = as.integer(substr(curr_date,6,7))
  Rate[i] = currUSD[(currUSD$Year == year) & (currUSD$Month == month),currency]
}

data = cbind(data, Rate)
rel_price = data$Price / data$Rate
data = cbind(data, rel_price)
head(data)
```

### Data Expository Graphs
```{r}
alive_artist = data[data$dead==0,]
dead_artist = data[data$dead==1,]
alive_mean = aggregate(alive_artist$rel_price, list(alive_artist$age), mean)
dead_mean = aggregate(dead_artist$rel_price, list(dead_artist$age), mean)
alive_age_division = cut(alive_mean$Group.1, breaks=seq(20,110,10))
dead_age_division = cut(dead_mean$Group.1, breaks=seq(20,110,10))
alive_mean = cbind(alive_mean, alive_age_division)
dead_mean = cbind(dead_mean, dead_age_division)

dead_mean = aggregate(dead_mean$x, list(dead_mean$dead_age_division), mean)
alive_mean = aggregate(alive_mean$x, list(alive_mean$alive_age_division), mean)

dead_mean$dead = rep(1,dim(dead_mean)[1])
alive_mean$dead = rep(0,dim(alive_mean)[1])

grouped_mean = rbind(dead_mean, alive_mean)
grouped_mean$dead = as.factor(grouped_mean$dead)
```

```{r}
gg <- ggplot(grouped_mean, aes(x=Group.1, y=x, fill=dead)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme(axis.text.x = element_text(angle=60,hjust=1)) +
  labs(title="Adj. Mean Painting Price", x="Age",y="Price")
plot(gg)
ggsave("Adj.Mean Painting.png",device="png", width=7, height=4.5)
```

```{r}

alive_geomrise = aggregate(alive_artist$geomrise, list(alive_artist$age), mean)
dead_geomrise = aggregate(dead_artist$geomrise, list(dead_artist$age), mean)
alive_age_division = cut(alive_geomrise$Group.1, breaks=seq(20,110,10))
dead_age_division = cut(dead_geomrise$Group.1, breaks=seq(20,110,10))
alive_geomrise = cbind(alive_geomrise, alive_age_division)
dead_geomrise = cbind(dead_geomrise, dead_age_division)

dead_geomrise = aggregate(dead_geomrise$x, list(dead_geomrise$dead_age_division), mean)
alive_geomrise = aggregate(alive_geomrise$x, list(alive_geomrise$alive_age_division), mean)

dead_geomrise$dead = rep(1,dim(dead_geomrise)[1])
alive_geomrise$dead = rep(0,dim(alive_geomrise)[1])

grouped_geomrise = rbind(dead_geomrise, alive_geomrise)
grouped_geomrise$dead = as.factor(grouped_geomrise$dead)
```


```{r}
gg <- ggplot(grouped_geomrise, aes(x=Group.1, y=x, fill=dead)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme(axis.text.x = element_text(angle=60,hjust=1)) +
  labs(title="Adj. Geomrise Painting Price", x="Age",y="Price")
plot(gg)
ggsave("Adj.Geomrise Painting.png",device="png", width=7, height=4.5)
```

```{r}
count_byartist = data %>% group_by(Artist,dead) %>% count() %>% arrange(desc(n))
count_byartist = count_byartist[1:20,]
count_byartist$Artist = factor(count_byartist$Artist, levels=count_byartist$Artist)
gg = ggplot(count_byartist, aes(x=Artist, y=n, fill=dead)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle=60,hjust=1)) +
  labs(title="Top Traded Artists (2010~2019)", x="Artist",y="Count")
plot(gg)
ggsave("Top Traded Artists.png",device="png", width=7, height=4.5)
```

```{r}
data[data$Artist %in% count_byartist$Artist,] %>% group_by(Artist, Year) %>% summarise(mean_price = mean(Price))
```

```{r}
# Histogram of response distribution
gg <- ggplot(data, aes(x=geomrise)) +
      geom_histogram() +
      labs(title="Response Distribution")
plot(gg)
```

```{r}
#What is the best cut of the age?
gg <- ggplot(data, aes(x=age, y=geomrise)) + 
  geom_point(aes(col=dead)) + 
  geom_smooth(method="loess", se=F) +
  labs(subtitle="Age vs Rise",
       y = "Geomrise",
       x = "Age",
       title = "Art Price",
       caption ="Source: self")
plot(gg)
```

```{r}
# Histogram of age distribution
gg <- ggplot(data, aes(x=age, fill=dead, color=dead)) + 
        geom_histogram( alpha=0.5, position = "identity") +
        labs(subtitle="Age Range Histogram")
plot(gg)
ggsave("Age Range.png",gg, device="png", width=7, height=4.5)
```

```{r}
# Investigate the variance of the age group through time
var_by_young = as.data.frame(data %>% group_by(young, Year) %>% summarise(var = var(log(Price))))
gg1 <- ggplot(var_by_young, aes(x=Year, y=var, group=young, color=young)) +
        geom_line() +
        theme(legend.position="none") + 
        labs(title="Yearly var. grouped by age-group")

range_by_young = as.data.frame(data %>% group_by(young, Year) %>% summarise(range = var(log((High-Low)/sqrt(Low*High)))))
gg2 <- ggplot(range_by_young, aes(x=Year, y=range, group=young, color=young)) +
        geom_line() +
        labs(title="Yearly range. grouped by age-group")

gg12 = arrangeGrob(gg1, gg2, ncol=2)
ggsave("Yearly Var.png",gg12, device="png", width=7, height=4.5)
```

- How price range effects the chance of beating or underperforming the expectation
- average prices over age group?


# DATA SUBSET

```{r}
#Subsetting the data with necessary variable - USING GEOMRISE
subdata = data[,c(12,2,13,14,9,11,5,6,7)]
# Death measured at the time of the selling
# So some artists indicated alive might be dead now!
levels(subdata$dead) = c(0,5)
colnames(subdata) <- c("GEOMRISE", "NAME", "YEAR","AGE", "LOC", "PRICE","LOW", "HIGH", "YOUNG")
head(subdata)
# Note that factor baseline is AMSTERDAM and YOUNG 0
# YOUNG =4 means dead
```

```{r}
subdata = subdata %>% arrange(NAME)
subdata = subdata %>% mutate(ID = as.numeric(group_indices(.,NAME)))

temp = tabulate(subdata$ID)
ID5 = which(temp >= 2)
#Dataset containing only artists with at least 5 sales
art5 = subdata[which(subdata$ID %in% ID5),]
#Make a new data frame with columns for ID, YOUNG, and Var(logrise)
varlogrise = rep(0, length(unique(art5$ID)))
young = varlogrise
count = 1
for (i in unique(art5$ID)) {
  temp = art5[which(art5$ID == i),]
  young[count] = as.numeric(as.character(temp$YOUNG[1]))
  varlogrise[count] = var(temp$GEOMRISE)
  count = count+1
}
#Make plot. Remember that 0 means dead, 1-5 is increasing in age.
boxplot(varlogrise ~ young, ylim=c(0,0.5))
```

### Data Modeling

```{r}
#Simple model only with fixed effects - GEOMRISE
linear.art = lm(GEOMRISE ~ YEAR + YOUNG + LOC + log((HIGH - LOW) / sqrt(LOW * HIGH)), data=subdata)
summary(linear.art)
qqnorm(residuals(linear.art))
qqline(residuals(linear.art))
mixed.art = lmer(GEOMRISE ~ (1|NAME) + YOUNG + YEAR + LOC + log((HIGH - LOW) / sqrt(LOW * HIGH)), data=subdata)
summary(mixed.art)
```

```{r}
#Now Subset the Data using the NOMINAL PRICE
subdata = data[,c(5,6,7,2,12,9,11,14,15,13)]
#subdata$young = as.numeric(as.character(subdata$dead))
#subdata[subdata$young == 0]$young = as.numeric((factor(cut(subdata[subdata$dead==0]$age, 4), labels=c(1,2,3,4))))
#subdata$young = subdata$young - 1
#subdata$young = as.factor(subdata$young)
# young = 0 indicates death, 1~5 are dividedcx subgroups
# Death measured at the time of the selling
# So some artists indicated alive might be dead now!
colnames(subdata) <- c("PRICE", "LOW", "HIGH", "NAME", "YEAR", "LOC","RATE", "YOUNG", "DEAD", "AGE")
subdata$REGION = subdata$LOC
subdata[subdata$LOC %in% c("London","Milan","Zurich","Amsterdam", "Paris"),]$REGION = "EU"
subdata[subdata$LOC %in% c("Dubai","Mumbai"),]$REGION = "WA"
subdata[subdata$LOC %in% c("Shanghai","Hong Kong"),]$REGION = "CN"
subdata[subdata$LOC %in% c("New York"),]$REGION = "US"
subdata$REGION = as.factor(subdata$REGION)
head(subdata)
```

```{r}
#Fitting Regression using nominal price
linear.art = lm(log(PRICE) ~ YEAR + YOUNG + LOC + I(log(HIGH*LOW)/2) + log(HIGH/LOW), data=subdata)
mixed.art = lmer(log(PRICE) ~ (1|NAME) + YOUNG + YEAR + LOC + I(log(HIGH*LOW)/2) + log(HIGH/LOW), data=subdata)
lrtstat = as.numeric(2*(logLik(mixed.art) - logLik(linear.art)))
pval = pchisq(lrtstat, 1, lower=FALSE)
data.frame(lrtstat, pval)
```

```{r}
# Exact  LR test to check random effects significance
exactLRT(mixed.art, linear.art)
```

```{r}
# Model Diagnositc
p2 = ggplot(subdata, aes(sample=log(PRICE))) + stat_qq() + stat_qq_line(color="red") + 
  ggtitle("QQplot") + theme_bw()
# p1 = ggplot(mixed.art, aes(.fitted, .resid))+geom_point()
#     p1<-p1+stat_smooth(method="loess")+geom_hline(yintercept=0, col="red", linetype="dashed")
#     p1<-p1+xlab("Fitted values")+ylab("Residuals")
#     p1<-p1+ggtitle("Residual vs Fitted Plot")+theme_bw()

# ggsave("./graphs/Model2_residplot.png",p1, device="png", width=7, height=4.5)
ggsave("./graphs/Model2_qqplot.png",p2, device="png", width=7, height=4.5)
```

```{r}
# Grouping locational covariates
subdata$REGION = subdata$LOC
subdata[subdata$LOC %in% c("London","Milan","Zurich","Amsterdam", "Paris"),]$REGION = "EU"
subdata[subdata$LOC %in% c("Dubai","Mumbai"),]$REGION = "WA"
subdata[subdata$LOC %in% c("Shanghai","Hong Kong"),]$REGION = "CN"
subdata[subdata$LOC %in% c("New York"),]$REGION = "US"
subdata$REGION = as.factor(subdata$REGION)
linear_regional.art = lm(log(PRICE) ~ YEAR + YOUNG + REGION + I(log(HIGH*LOW)/2) + log(HIGH/LOW), data=subdata)
mixed_regional.art1 = lmer(log(PRICE) ~ (1|NAME) + AGE + YEAR + REGION + I(log(HIGH*LOW)/2) + log(HIGH/LOW), data=subdata)
mixed_regional.art2 = lmer(log(PRICE) ~ (1 + DEAD|NAME)  + YOUNG + YEAR + REGION + I(log(HIGH*LOW)/2) + log(HIGH/LOW), data=subdata)
#summary(mixed_regional.art)
```


```{r}
# Model Diagnositc
p2 <- ggplot(subdata, aes(sample=log(PRICE))) + stat_qq() + stat_qq_line(color="red") + 
  ggtitle("QQ Plot") +
  theme_bw()
p1 <- ggplot(mixed_regional.art1, aes(.fitted, .resid))+geom_point()
    p1<-p1+stat_smooth(method="loess")+geom_hline(yintercept=0, col="red", linetype="dashed")
    p1<-p1+xlab("Fitted values")+ylab("Residuals")
    p1<-p1+ggtitle("Residual vs Fitted Plot")+theme_bw()

ggsave("./graphs/Inference_model2",grid.arrange(p2, p1, ncol=2), device="png", width=7, height=4.5)
```

### Filtering out for dead artists
```{r}
# Analyzing artists who have passed away during the time frame
## subset the artist who died between 2010~2019
artist_passed_1119 = unique(data[data$Death %in% seq(2011,2019),]$Artist)
## from those artist, subset the artist who have sales record for both before and after
subdata2 = data[data$Artist %in% artist_passed_1119,]
sales_count = subdata2 %>% count(Artist) 
# Subset artist who have more than 1 sales record
sales_before_dead = subdata2[is.na(subdata2$Death),] %>% count(Artist)
sales_after_dead = subdata2[!is.na(subdata2$Death),] %>% count(Artist)
both_artist = intersect(sales_before_dead$Artist, sales_after_dead$Artist)

subdata2 = subdata2[subdata2$Artist %in% both_artist,]

subdata2 = subdata2[,c(5,6,7,2,12,9,11,13,15)]
colnames(subdata2) <- c("PRICE", "LOW", "HIGH", "NAME", "YEAR", "LOC","RATE", "AGE", "DEAD")
subdata2$REGION = subdata2$LOC
subdata2[subdata2$LOC %in% c("London","Milan","Zurich","Amsterdam", "Paris"),]$REGION = "EU"
subdata2[subdata2$LOC %in% c("Dubai","Mumbai"),]$REGION = "WA"
subdata2[subdata2$LOC %in% c("Shanghai","Hong Kong"),]$REGION = "CN"
subdata2[subdata2$LOC %in% c("New York"),]$REGION = "US"
subdata2$REGION = as.factor(subdata2$REGION)
head(subdata2)
```

```{r}
mm1 = lmer(log(PRICE) ~ (1|NAME) + DEAD + I(log(HIGH*LOW)/2) + log(HIGH/LOW) + REGION + YEAR , data=subdata2)
mm2 = lmer(log(PRICE) ~ (1 + DEAD|NAME) + DEAD + I(log(HIGH*LOW)/2) + log(HIGH/LOW) + REGION + YEAR , data=subdata2)
anova(mm1,mm2)
```

```{r}
# Model Diagnositc
p2 <- ggplot(subdata2, aes(sample=log(PRICE))) + stat_qq() + stat_qq_line(color="red") + 
  ggtitle("QQ Plot") +
  theme_bw()
p1 <- ggplot(mm2, aes(.fitted, .resid))+geom_point()
    p1<-p1+stat_smooth(method="loess")+geom_hline(yintercept=0, col="red", linetype="dashed")
    p1<-p1+xlab("Fitted values")+ylab("Residuals")
    p1<-p1+ggtitle("Residual vs Fitted Plot")+theme_bw()

ggsave("./graphs/Inference_model3",grid.arrange(p2, p1, ncol=2), device="png", width=7, height=4.5)
```




### Variance modeling
```{r}
var_by_young = as.data.frame(data %>% group_by(young, Year) %>% summarise(var = var(geomrise)))
range_by_young = as.data.frame(data %>% group_by(young, Year) %>% summarise(range = var(log((High-Low)/sqrt(Low*High)))))
var_data = merge(var_by_young, range_by_young)
var_linear = lm(var ~ range + Year + young, data=var_data)
range_linear = lm(range ~ Year + young, data=var_data)
summary(range_linear)
```


```{r}
#Try just on Art5
#Create ID
art5 = subdata %>% arrange(NAME)
art5 = art5 %>% mutate(ID = as.numeric(group_indices(.,NAME)))
#create inputs for function
y = log(art5$PRICE)
id = art5$ID
X = unname(model.matrix(~REGION + YOUNG + YEAR + I(log(HIGH*LOW)/2) + log(HIGH/LOW), data = art5))
U = unname(model.matrix(~1, data = art5))
Z = unname(model.matrix(~DEAD, data = art5))
nIter = 10
nBoot = 100
#Bootstrap confidence intervals
set.seed(19)
art5ResBoot = algorithm3BootInt(y, X, U, Z, id, nIter, nBoot, printN = FALSE)
#Model-based standard errors
art5Res = algorithm3SE(y, X, U, Z, id, nIter, getSigmaSE = 1)
sigma5SE = art5Res[[10]]
sigma5SE = sigma5SE[1]
```

Additional Analysis
+ Variability of ranges for younger artist is higher??? Thus analyst see this as uncertainty in price? How fast does range stabilization happpen for an art product?
+ Logisitic model that allows prediction for sales over or 

#### EXTRACT

```{r}
#model = lmer(log(PRICE) ~ (1|NAME) + YOUNG + YEAR + REGION + I(log(HIGH*LOW)/2) + log(HIGH/LOW), data=subdata)
model = lmer(log(PRICE) ~ (1 + DEAD|NAME) + DEAD + I(log(HIGH*LOW)/2) + log(HIGH/LOW) + REGION + YEAR , data=subdata2)
res.table = as.data.frame(coef(summary(model)))
ran.table = as.data.frame(VarCorr(model))
print(xtable(res.table, type = "latex", digits=4), file = "./latex_table/coef.tex")
print(xtable(ran.table, type = "latex", digits=4), file = "./latex_table/var.tex")
```


